\documentclass[onecolumn, amsmath, amsfonts, amssymb]{aastex62}
\usepackage{mathtools}
\usepackage{natbib}
\usepackage{bm}
\newcommand{\vdag}{(v)^\dagger}
\newcommand\aastex{AAS\TeX}
\newcommand\latex{La\TeX}


\newcommand{\Div}[1]{\ensuremath{\nabla\cdot\left( #1\right)}}
\newcommand{\DivU}{\ensuremath{\nabla\cdot\bm{u}}}
\newcommand{\angles}[1]{\ensuremath{\left\langle #1 \right\rangle}}
\newcommand{\KS}[1]{\ensuremath{\text{KS}(#1)}}
\newcommand{\KSstat}[1]{\ensuremath{\overline{\text{KS}(#1)}}}
\newcommand{\grad}{\ensuremath{\nabla}}
\newcommand{\RB}{Rayleigh-B\'{e}nard }
\newcommand{\stressT}{\ensuremath{\bm{\bar{\bar{\Pi}}}}}
\newcommand{\lilstressT}{\ensuremath{\bm{\bar{\bar{\sigma}}}}}
\newcommand{\nrho}{\ensuremath{n_{\rho}}}
\newcommand{\approptoinn}[2]{\mathrel{\vcenter{
	\offinterlineskip\halign{\hfil$##$\cr
	#1\propto\cr\noalign{\kern2pt}#1\sim\cr\noalign{\kern-2pt}}}}}

\newcommand{\appropto}{\mathpalette\approptoinn\relax}
\newcommand{\pro}{\ensuremath{\text{Ro}_{\text{p}}}}
\newcommand{\con}{\ensuremath{\text{Ro}_{\text{c}}}}

\usepackage{color}
\newcommand{\gv}[1]{{\color{blue} #1}}

%% Tells LaTeX to search for image files in the 
%% current directory as well as in the figures/ folder.
\graphicspath{{./}{figs/}}

\begin{document}
\section{State of the problem}
In these thermal simulations, as far as I can tell, we have \emph{eight} variables
\begin{enumerate}
\item $B$, the integrated total entropy leading to buoyancy
\item $V$, the thermal volume
\item $R$, the radius of the thermal
\item $h$, the thermal's vertical extent (`radius' in z-direction)
\item $w$, the thermal bulk vertical velocity
\item $z$, the height of the thermal
\item $\Gamma$, the circulation of the vortex ring
\item $\rho$, the local density at the height of the thermal.
\end{enumerate}
From the simulations we've done so far, when viscous heating is neglected, we have the
following understanding of thermals:
\begin{enumerate}
\item $B = \text{const} \propto \rho s_1$
\item $h = A R$, where $A$ is a constant [this assumption may not be true at high stratification].
\item $V = (4/3)\pi h R^2 = (4/3)\pi A R^3 = V_0 R^3$.
\item $w = \frac{\partial z}{\partial t} = -\frac{\partial d}{\partial t}$ (with depth $d \equiv L_z - z$)
\item $\rho = (1 + L_z - z)^{m_{ad}} = (1 + d)^{m_{ad}}$, with $m_{ad} = (\gamma-1)^{-1} = 1.5$ (for our polytropes)
\item $\rho V w \sim B t$ (momentum linearly increases)
\item $\rho \pi R^2 \Gamma \sim B t$ \citep[e.g., eqn 36 of ][]{shivamoggi2010}.
\end{enumerate}
That's 8 unknowns, and 7 equations. There's one piece of the puzzle we're missing:
what happens with $\Gamma$ over time?
Ideally, $\Gamma = \text{const}$, but we can't be sure that's the case. 

\section{Low stratification limit}
In the limit of a lightly stratified thermal, we make one more assumption:
\begin{enumerate}
\item $\Gamma = \text{const}$ (our eqn \# 8)
\end{enumerate}
We also know that, under this assumption, and also in this regime where eqn 2 is valid,
from dividing eqn 6 by 7, 
$$
w R \sim \frac{\pi\Gamma}{V_0}.
$$
In general, this suggests that $w \propto t^{-\alpha}$ and $R \propto t^{\alpha}$, such that
when multiplied together they return a constant. With this, we return to eqn 6 or 7, and get
\begin{equation}
\rho R^2 = \frac{B}{\pi \Gamma} t
\label{eqn:theory_eqn_6_and_7}
\end{equation}
With our ansatz of $R \propto t^{\alpha}$, we know that $\rho \sim t^{1 - 2\alpha}$.
But we also know from the stratification of our domain that 
$\rho = (1 + d)^{m_{ad}} \sim t^{1 - 2\alpha}$, and that $d \sim -t^{1-\alpha}$ (from
theory equation 4, definitionally). If we neglect the `1' in the $\rho$ definition for now,
we thus get $m_{ad}(1 - \alpha) = (1 - 2\alpha)$, or
$$
\alpha = \frac{1 - m_{ad}}{2 - m_{ad}} = -1
$$
><

\section{What do we need to do in the stratified limit?}
Figure out what happens to the circulation.


\bibliography{biblio.bib}
\end{document}
